##Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website
here: http://groupware.les.inf.puc-rio.br/har (http://groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset).

## PREPROCESSING

### Load the train and quiz data

We load both datasets from their files, treating "#DIV/0!" and "NA" values as NA:

```{r}
train <- read.csv("pml-training.csv", header=TRUE, sep=",", quote = "\"", na.strings=c("#DIV/0!","NA") )
quiz <- read.csv("pml-testing.csv", header=TRUE, sep=",", quote = "\"", na.strings=c("#DIV/0!","NA") )
```

Having `r nrow(train)` rows and `r ncol(train)` attributes in **train** dataset, 
and `r nrow(quiz)` rows and `r ncol(train)` attributes in **quiz** dataset to predict.

### Remove the id attributes

Id attributes will only cause overfitting in our training model.
We remove them from both train and quiz datasets:

```{r}
train$X <- NULL; quiz$X <- NULL;
train$raw_timestamp_part_1 <- NULL; quiz$raw_timestamp_part_1 <- NULL
train$raw_timestamp_part_2 <- NULL; quiz$raw_timestamp_part_2 <- NULL
train$cvtd_timestamp <- NULL; quiz$cvtd_timestamp <- NULL
quiz$problem_id<-NULL
```

Having `r ncol(train)` attributes in **train** dataset, 
and `r ncol(train)` attributes in **quiz** dataset to predict.

### Remove the empty (or almost) attributes

Look for empty, or almost empty, attributes in the quiz dataset:

```{r}
columnasPobladas <- NULL; columnasDespobladas <- NULL
filas<-dim(quiz)[1]; 
for(i in c(1:ncol(quiz))) {
  if (sum(is.na(quiz[,i])) / filas >= .9) { 
		columnasDespobladas <- c(columnasDespobladas, colnames(quiz[i]))
	}
	else {
		columnasPobladas <- c(columnasPobladas, colnames(quiz[i]))
	}
}
```

Having `r length(columnasPobladas)` populated attributes, and `r length(columnasDespobladas)` empty attributes.

Then, remove them from both train and quiz datasets:

```{r}
for (columna in columnasDespobladas) {
	train[,columna] <- NULL
  quiz[,columna] <- NULL
}
```

Now, we have in the **train** dataset:

```{r, echo=FALSE}
table(sapply(train[1,],class))
```

And in the **quiz** dataset, the same except for the outcome factor:

```{r, echo=FALSE}
table(sapply(quiz[1,],class))
```

### Convert non-factor attributes to numeric

To deal only with numeric types, avoiding integer:

```{r}
colToConvert <- sapply(train, class)!="factor"
for(i in c(1:ncol(train))) {
  if (colToConvert[i]) {
    train[,i] <- as.numeric(train[,i])
  }
}

colToConvert <- sapply(quiz, class)!="factor"
for(i in c(1:ncol(quiz))) {
  if (colToConvert[i]) {
    quiz[,i] <- as.numeric(quiz[,i])
  }
}
```

Now, we have in the **train** dataset:

```{r, echo=FALSE}
table(sapply(train[1,],class))
```

And in the **quiz** dataset, the same except for the outcome factor:

```{r, echo=FALSE}
table(sapply(quiz[1,],class))
```

## CLASSIFICATION MODEL

```{r, echo=FALSE}
require(caret); require(randomForest)
```

### Data splitting

We split the train dataset into a 70% for **training**, and 30% for **testing**:

```{r}
inTrain <- createDataPartition(y=train$class, p=0.7, list=FALSE)
training <- train[inTrain,]
testing <- train[-inTrain,]
```

Having `r nrow(training)` rows to **training** the model, and `r nrow(testing)` rows for **testing** the model.

### Training

We train a classification model using Random Forest, [that doesn't need for for cross-validation or a separate test set to get an unbiased estimate of the test set error](http://www.stat.berkeley.edu/~breiman/RandomForests/cc_home.htm#ooberr):

```{r}
set.seed(32343)
modelFitRF <- train(classe~.,data=training, method="rf")
```

```{r, echo=FALSE}
modelFitRF
```

Obtaining this model:

```{r, echo=FALSE}
modelFitRF$finalModel
```

### Evaluating the model on the testing dataset

```{r}
confusionMatrix(testing$classe,predict(modelFitRF,testing))
```

### Predicting

And, finally, applying this model to the quiz dataset, we make our predictions:

```{r}
predict(modelFitRF,quiz)
```
